---
title: "Introduction-excercise"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction excercise

The objective of this excercise is to be familiar with data sourcing and basic data manipulation for both R and python. After this excercise you should be able to load files, create data frames and understand basic data transformation. If you are already familiar with data manipulation you might want to skip this excercise.


### Packages & libraries

In R Packages are the fundamental units of reproducible code. They include reusable R functions, the documentation that describes how to use them, and sample data. Packages enhance R by improving existing base R functionalities, or by adding new ones. There are general packages like dplyr and data.table that help with data cleaning and wrangling to more advatages pages such as "NLP", that allow text processing for several languges

Installing packages can be done with the following command
```{r eval=FALSE}
install.packages("data.table")

```
"data.table" is the name of the package that we are installing onto R, this will allow us to have access to new functions that are not present in base R such as, subset, group, etc.

In order to attach this packages to our current R session we can use two commands

```{r eval=FALSE}
library(data.table)
require(data.table)

```
Both commands when used in R will give you access to the functions contained inside the package, the only difference is that library() will producean error if the package is not installed, and require() will only give a warninig.

For, example if we want to run python code in R we can use the package "reticulate"

```{r  message = FALSE, eval=TRUE}
#install.packages("reticulate")
library(reticulate)

#knitr::knit_engines$set(python = reticulate::use_python('C:/tools/anaconda3/5.3.0'))

```
In this case we first install the package and then attach it to our session with the library() command this will allow us to run the following code, which would have been impossible in base R.(Note: since reticulate was already installed in my machine I used "#" to comment this code so it does not run and reinstall the package)


```{python eval=FALSE}
#pip install numpy
#conda inöstall numpy

import pandas as pd
import sys
print(print(sys.version))


```
* In python packages can be installed wiht pip or using Anaconda in the command prompt

### Data sourcing

The most importan part of a data scientist job is the data. Logically, this means that the first step into learning about data science is not on statistics on comple mathematics but understanding how obtain data from different sources.
This excercises will help you to load data into your prefered work enviroment.

#### Reading from files

```{r eval=FALSE}
require(data.table) # general purpose package for data manipulation
require(readxl) # package to read data from excel files

# If you want to choose the file from a gui you can use
guidata <- read_excel(file.choose())

# if you want to read data from a copied excel table on clipboard
clipboarddata <- read.table(file = "clipboard", sep = "\t", header=TRUE)


```


If you want to read files from other directories without having to open a file choosing GUI
```{r}

require(data.table) # general purpose package for data manipulation
require(readxl) # package to read data from excel files

txtdata<-read.table('biostats.txt',header = TRUE,sep = ',')
exceldata<-read_excel("biostats.xlsx")
csvdata<-read.table('biostats.csv',header = TRUE,sep = ',')

summary(csvdata)

```
In this case the data is on the work directory, but if you want to use data from other directories this can be done by specifying the file path as the first argument of the function
e.g. 
```{r eval=FALSE}
data<-read.table('C:/USER/DATA/table.csv',header=FALSE,sep = ',')
```
```{r}

```

In a similar manner you can export data from R to a file
```{r eval=FALSE}
require(data.table)

write.table(csvdata, 'example2.txt', append = FALSE, sep = " ", dec = ".",row.names = TRUE, col.names = TRUE)
```


### Basic data manipulation

once you have access to your data, then the real work can start. in this brief excercises we will get an overview on how to use R to manipulate data in a manner that is not possible in excel.

First let's load some data and let's print the first rows

```{r eval=TRUE}
require(tidyr)

my_data <- as.data.frame(iris)

head(my_data)

```
Now lets's reorganize the data according to one variable and look at the table
```{r }

# Rearrange data
my_data2<-my_data %>% arrange(Sepal.Length)
head(my_data2)
```
Now let's try to other types basic manipulation like change the functions

```{r eval=TRUE}


# change column values

my_data2<-my_data %>% mutate_all(funs(./2.54))
head(my_data2)

```
As you can see we try to divide a non valid data type like 'text' we obtain 'NA', to avoid this we can select or add new columns

```{r}

# Transform variables

my_data2 <- transform(my_data, logLength=log(Sepal.Length), logWidth=log(Sepal.Width))

head(my_data2)

```

To close up let's do basic data merge

```{r}



# add new data
allData <- rbind(csvdata, txtdata)

# add new variables

allData2 <- cbind(my_data[1:18,], exceldata)

head(allData)

head(allData2)
```

