---
title: "Tabular data classification"
output: html_notebook
---

## Getting data
A classification task requires two things a label and the observation corresponding to that label, In this case we will use a dataset with description of different mushrooms, and we will try to classify if they are edible or poisonus


```{r}
library(tidyverse)
mushrooms<-read.csv('mushrooms.csv')
head(mushrooms)
```
```{r}
summary(mushrooms)
```

It seems like we dont have any obvious data quality issues, let's take a closer look at our label variable
which is the class column, its distributions is as follows

```{r}
table(mushrooms$class)
```

In this case the data labels are very well balanced so random forest is ideal for this kind of data, but before classifying let's further explore

```{r}
library(ggplot2)

p = ggplot(mushrooms,aes(x=odor,  
                    y=spore.print.color, 
                    color=class))

p + geom_jitter(alpha=0.3) +  
  scale_color_manual(breaks = c('e','p'),
                     values=c('darkgreen','red'))

```
In this case:
odor (almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s)
spore-print-color(black=k,brown=n,buff=b,chocolate=h,green=r,orange=o,purple=u,white=w,yellow=y)

Data seems like there is enough difference between the observersion of edible mushrooms and poisonus, let's prepare the data for classification
```{r}
library(randomForest)
library(caret)
library(caTools)

set.seed(2012)
mushrooms<-mushrooms[sample(nrow(spam)),]

set.seed(123)   
samplem <- sample.split(mushrooms$class,SplitRatio = 0.75) 
mush.train <-subset(mushrooms,samplem ==TRUE) 
mush.test<-subset(mushrooms, samplem ==FALSE)



```

Let's look at the class balance for the training set

```{r}
table(mush.train$class)/nrow(mush.train)  
```
And now for the original dataset

```{r}
table(mushrooms$class)/nrow(mushrooms)  
```

So this function split help us keep the same balance for training and testing as in our original dataset

```{r}
rf = randomForest(class ~ .,  
                   ntree = 100,
                   data = mush.train)
plot(rf)  
```

Now let's check what are the more important variables

```{r}
varImpPlot(rf,  
           sort = T,
           n.var=10,
           main="Top 10 - Variable Importance")
```

Now lets evaluate our model

```{r}
mpred<-predict(rf ,mush.test)
confusionMatrix(table(predicted=mpred,actual=mush.test$class), mode = "everything")

```

