---
title: "Data cleaning"
output: html_notebook
---

## Data quality

```{r}
data<-read.csv('Water_Right_Applications.csv',stringsAsFactors = FALSE)
head(data)
```

This dataset has several issues, however opening individual files or databases is quite time consuming, so let's create a function to summarise different aspects of the datasets

```{r}
dfsummary<-function(df){
 
  summary=data.frame()
  for(i in 1:ncol(df)){
    
    if(length(df[!(is.na(df[,i])),i])==0){
      
      temp=data.frame(counts=0,
                      uniques=0,
                      missing=nrow(df[(is.na(df[,i])),i]),
                      miss_per=(1),
                      percent_mfv=(0)) 
      
      
    }else{
      
      temp=data.frame(counts=length(df[!(is.na(df[,i])),i]),
                      uniques=uniqueN(df[,i]),
                      missing=length(df[(is.na(df[,i])),i]),
                      data_type=class(data[,i]),
                      no_space=length(data[data[,i]==' ',i]))
      
    }
    
    
    rownames(temp)<-colnames(df)[i]
    
    summary<-rbind(summary,temp)
    
    #print(i)
  }
  
  return(summary)
  
}

data_summary<-dfsummary(data)

data_summary
```

Let's try to fix the most obvious errors first let's drop unecessary data

```{r}
data<-data[,-25]
colnames(data)
```

Next we want to fix dates in a format that is sortable

```{r}
library(tidyverse)
library(lubridate)

data$date <- mdy(data$PRIORITY_DATE)
data$weekday <- wday(data$date,label = TRUE,locale = 'English')

data %>% filter(weekday=='Wed') %>% head()
```

What to do with empty spaces and empaty values is up to the application. However is a good to change all  data with empty ispaces into empty values
```{r}
print(nrow(data[is.na(data[,19]),]))
print(nrow(data[(data[,19]==' '),]))

data[(data[,19]==' '),19]<-NA
print(nrow(data[is.na(data[,19]),]))
```




